### 2020-04-27 Organismal_paper/Organismal_manuscript_and_code.Rmd
progress in organismal paper

```{r setup, include=FALSE}
	# Load packages
	## General
library(knitr)
library(tidyverse)
library(stringr)
library(reshape2)
	## Biological
library(Rphylip)
library(arbutus)
library(vegan)
library(ape)
library(phangorn)
library(phytools)
library(picante)
library(geiger)
library(phylobase)
library(fields)
library(phylosignal)
library(geomorph)
library(phylopath)
library(phylolm)
	## Graphics
library(FactoMineR)
library(factoextra)
library(corrplot)
library(BAMMtools)
library(gridExtra)
library(xtable)
library(colorRamps)

@@

In [@damianserrano2020shaped], we collected the most extensive morphological dataset on siphonophore tentilla and nematocysts using state-of-the-art microscopy techniques, and expanded the taxon sampling of the phylogeny to disentangle the evolutionary history. The analyses we carried out led to novel insights into the evolution of predatory specialization. However, this work left behind many fascinating discoveries on the evolutionary history of tentilla morphology, including nematocyst novelties, phenotypic integration, convergent evolution, paedomorphosis, and predictions on diet using morphology. Here, we present these findings and shed light on the natural history of siphonophores.

@@

We explore the evolutionary history and diversification of tentilla morphologies using phylogenetic comparative methods on the morphological dataset and phylogeny published in [@damianserrano2020shaped].  We fitted different models generating the observed data distribution given the phylogeny for each continuous character using the function fitContinuous in the R package *geiger* [@harmon2007geiger]. These models include a non-phylogenetic whie-noise model (WN), a neutral divergence Brownian Motion model (BM), an early-burst decreasing rate model (EB), and an Ornstein-Uhlenbeck model with stabilizing selection around a fitted optimum trait value. We then ranked the models in order of increasing parametric complexity (WN, BM, EB, OU), and compared the corrected Akaike Information Criterion (AICc) support scores [@sugiura1978further] to the lowest (best) score, using a cutoff of 2 units to determine significantly better support. When the best fitting model was not significantly better than a less complex alternative, we selected the least complex model (SM13). We calculated model adequacy scores using the R package *arbutus* [@pennell2015model] (SM14). We calculated phylogenetic signals in each of the measured characters using Blomberg’s K [@blomberg2003testing] (SM13).

@@

*Evolutionary history of tentillum morphology* – [@damianserrano2020shaped] produced the most speciose siphonophore molecular phylogeny to date, while incorporating the most recent findings in siphonophore deep node relationships. This phylogeny revealed for the first time that the genus *Erenna* is the sister to *Stephanomia amphytridis*. *Erenna* and *Stephanomia* bear the largest tentilla among all siphonophores, thus their monophyly indicates that there was a single evolutionary transition to giant tentilla. Siphonophore tentilla range in size from ~30 µm in some *Cordagalma* specimens to 2-4 cm in *Erenna* species, and up to 8 cm in *Stephanomia amphytridis* [@pugh2014review]. Most siphonophore tentilla measure between 175 and 1007 µm (1st and 3rd quartiles), with a median of 373 µm. The extreme gain of tentillum size in this newly found clade may have important implications for access to large prey size classes.

@@

  The clades defined in [@damianserrano2020shaped] are characterized by unique evolutionary innovations in their tentilla. The clade Eucladophora (containing Pyrostephidae, Euphysonectae, and Calycophorae) encompasses all of the extant Siphonophore species (178 of 186) except Cystonects and *Apolemia*. Innovations that arose along the stem of this group include spatially segregated heteroneme and haploneme nematocysts, terminal filaments, and elastic strands (Fig. \@ref(figure7)). Pyrostephids evolved a unique bifurcation of the axial gastrovascular canal of the tentillum known as the "saccus" [@totton1965synopsis]. The stem to the clade Tendiculophora (clade containing Euphysonectae and Calycophorae) subsequently acquired further novelties such as the desmoneme and rhopaloneme (acrophore subtype ancestral) nematocysts on the terminal filament (Fig. \@ref(figure7)), which bears no other nematocyst type. These are arranged in sets of 2 parallel rhopalonemes for each single desmoneme [@skaer1988formation;@skaer1991remodelling]. The involucrum is an expansion of the epidermal layer that can cover part or all of the cnidoband (Fig. \@ref(figure2)). This structure, together with differentiated larval tentilla, appeared in the stem branch to Clade A physonects. Calycophorans evolved novelties such as larger desmonemes at the distal end of the cnidoband, pleated pedicles with a “hood” (here considered homologous to the involucrum) at the proximal end of the tentillum, anacrophore rhopalonemes, and microbasic mastigophore-type heteronemes. While calycophorans have diversified into most of the extant described siphonophore species (108 of 186), their tentilla have not undergone any major categorical gains or losses since their most recent common ancestor. Nonetheless, they have evolved a wide variation in nematocyst and cnidoband sizes.

@@

*Evolution of tentillum and nematocyst characters* – One-third of the characters measured in [@damianserrano2020shaped] support a non-phylogenetic generative model, indicating they are not phylogenetically conserved (SM\@ref(ModelSupport)). Most (74%) characters present a significant phylogenetic signal, yet only total nematocyst volume, haploneme length, and heteroneme-to-cnidoband length ratio had a phylogenetic signal with K larger than 1. Total nematocyst volume and cnidoband-to-heteroneme length ratio showed strongly conserved phylogenetic signals. The majority (67%) of characters support BM models, indicating a history of neutral constant divergence. No relationship between phylogenetic signal and BM model support was found. Haploneme nematocyst length is the only character with support for an EB model of decreasing rate of evolution with time. No character had support for a single-optimum OU model (when uninformed by feeding guild regime priors).

@@

![\label{figure10} Phylomorphospace showing haploneme and heteroneme elongation (log scaled). Orange area delimits rod-shaped haplonemes, the blue area covers oval and round-shaped haplonemes. Smaller dots and lines represent phylogenetic relationships and ancestral states of internal nodes under BM. Species nodes in red lack either haplonemes or heteronemes, and their values projected onto the axis of the nematocyst type they bear. Cystonects have no tentacle heteronemes and are projected onto the haploneme axis. Apolemiids have no tentacle haplonemes and are projected onto the heteroneme axis. ](~/tentilla_morph/Figures_final/figure10.jpg)

@@

*Generating dietary hypotheses using tentillum morphology* – For many siphonophore species, no feeding observations have yet been published. We generated hypotheses about the diets of these understudied siphonophores based on their known tentacle morphology and the associations between morphology and diet reported here for other species using the linear discriminant analysis of principal components (DAPC) fitted in [@damianserrano2020shaped]. This provides concrete predictions to be tested in future work and helps extrapolate our findings to many poorly known species that are extremely difficult to collect and observe.
The discriminant analysis in [@damianserrano2020shaped] for feeding guild (7 principal components, 4 discriminants) produced 100% discrimination, and the highest loading contributions were found for the characters (ordered from highest to lowest): Involucrum length, heteroneme volume, heteroneme number, total heteroneme volume, tentacle width, heteroneme length, total nematocyst volume, and heteroneme width. We used the predictions from this discriminant function to generate hypotheses about the feeding guild of 45 species in the morphological dataset. This extrapolation predicts that two other *Apolemia* species are gelatinous prey specialists like *Apolemia rubriversa*, and predicts that *Erenna laciniata*is a fish specialist like *Erenna richardi*. When predicting soft- and hard-bodied prey specialization, the DAPC achieved 90.9% discrimination success, only marginally confounding hard-bodied specialists with generalists (SM\@ref(DAPCsofthard)). The main characters driving this discrimination are involucrum length, heteroneme number, heteroneme volume, tentacle width, total nematocyst volume, total haploneme volume, elastic strand width, and heteroneme length. We only selected prey types with sufficient variation in the data to carry out these analyses (copepods, fish, and large crustaceans). While the presence of fish or large crustaceans in the diet cannot be unambiguously discriminated using tentillum morphology (SM\@ref(DAPCfish),SM\@ref(DAPCcrustacean)), specialization on fish or large crustacean prey can be fully disentangled (SM\@ref(DPACguilds)).

@@

*On the evolution of tentilla morphology* -- The evolutionary rate covariance results in [@damianserrano2020shaped] indicate that tentilla are not only phenotypically integrated but also show patterns of evolutionary modularity, where different sets of characters appear to evolve in stronger correlations among each other than with other characters [@wagner1996homologues]. This may be indicative of the underlying genetic and developmental dependencies among closely homologous nematocyst types (such as desmonemes and rhopalonemes) and structures. In addition, these evolutionary modules point to hypothetical functional modules. For example, the coiling degree of the cnidoband and the extent of the involucrum have correlated rates of evolution, while high-speed videos (pers. obs.) show that the involucrum helps direct the whiplash of the uncoiling cnidoband distally (towards the prey).
The clade Tendiculophora contains far more species than its relatives Cystonectae, Apolemiidae, and Pyrostephidae. An increase in clade richness and ecological diversification can be triggered by a ‘key innovation’ [@simpson1955major]. The evolutionary innovation of the Tendiculophora tentilla with shooting cnidobands and modular regions may have facilitated further dietary diversification.

Haploneme and heteroneme shape share 21% of their variance across extant values, and 53% of the variance in their shifts along the branches of the phylogeny. However, much of this correlation is due to the contrast between Pyrostephidae and their sister group Tendiculophora (Fig. \@ref(figure4)). We searched for rate regime shifts in the evolution of nematocyst shape characters using a Bayesian Analysis of Macroevolutionary Mixtures (BAMM) [@rabosky2014bamm] (SM32-35). BAMM located a regime shift in heteroneme shape evolution on the branches leading to *Agalma* and *Athorybia* (SM33). For the rates of haploneme shape evolution, BAMM identified two main independent regime shifts (Fig. \@ref(figure10)): one in the branch leading to Codonophora (anisorhizas diverging from cystonects’ spherical isorhizas), and one in the branch leading to Clade B physonects (SM35). Clade B includes *Erenna*, *Stephanomia*, *Marrus*, and rhodaliids. Most of these taxa have rod-shaped anisorhizas, but *Erenna* have oval ones. No significant regime shift patterns were identified in the evolution of desmoneme and rhopaloneme shape.

@@

Our work identifies a novel example of convergent evolution. The region of the tentillum morphospace (SM4) occupied by calycophorans was independently (and more recently) occupied by the physonect *Frillagalma vityazi*. Like calycophorans, *Frillagalma* tentilla have small C-shaped cnidobands with a few rows of anisorhizas. Unlike calycophorans, they lack paired elongate microbasic mastigophores. Instead, they bear exactly three oval stenoteles, and their cnidobands are followed by a branched vesicle, unique to this genus. Their tentillum morphology is very different from that of other related physonects, which tend to have long, coiled, cnidobands with many paired oval stenoteles. Most studies on calycophoran diets have reported their prey to be primarily composed of small crustaceans, such as copepods or ostracods [@purcell1981dietary;@purcell1984functions]. The diet of *Frillagalma vityazi* is unknown, but this morphological convergence suggests that they evolved to capture similar kinds of prey. The DAPCs in [@damianserrano2020shaped] predict that *Frillagalma* has a generalist niche with both soft and hard-bodied prey, including copepods.

@@

```{r code}
# Set paths to input data
setwd("~/tentilla_morph/Organismal_paper/")

@@

### COMPARATIVE ANALYSES ###

@@

## PCA ##
Q_sharedmean_logs = sharedmean_logs[,-1]
#Q_sharedmean_logs = castmean_logs[,-1]
#rownames(Q_sharedmean_logs) = castmean_logs$Species
rownames(Q_sharedmean_logs) = sharedmean_logs$Species
raw_matrix = Q_sharedmean_logs[,c(1:2,4:20)] %>% .[which(!is.na(rowSums(.))),]
raw_matrix_notf = Q_sharedmean_logs[,c(1:2,4:20)] %>% .[,c(1:7,12:17)] %>% .[which(!is.na(rowSums(.))),]
raw_matrix_NaZeroes = Q_sharedmean_logs[,c(1:2,4:20)]
raw_matrix_NaZeroes[is.na(raw_matrix_NaZeroes)]<-0
compound_matrix = Q_sharedmean_logs[which(!is.na(rowSums(Q_sharedmean_logs))),c(3,21:31)]

#Using simple characters
PCA(raw_matrix_NaZeroes) -> Pca_raw
PCA(raw_matrix_notf) -> Pca_raw
Pca_raw %>% fviz_contrib(choice="var", axes=1, sort.val="desc")
Pca_raw %>% fviz_contrib(choice="var", axes=2, sort.val="desc")
Pca_raw %>% fviz_pca_biplot( col.var="contrib", gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), repel = TRUE) #For figure A
Pca_raw %>% fviz_pca_biplot( col.var="contrib", gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), repel = TRUE, axes=c(3,4)) 
raw_tree = drop.tip(ultram, which(!(ultram$tip.label %in% rownames(raw_matrix_notf))))
phylomorphospace(tree=raw_tree, Pca_raw$ind$coord[,1:2], label = "horizontal", xlab = "PC1", ylab = "PC2") #For figure B
multiPhylosignal(Pca_raw$ind$coord, raw_tree)
physignal(Pca_raw$ind$coord, raw_tree)

@@

#PhyloPCA on simple character
PPCA_raw = phyl.pca(raw_tree, raw_matrix_notf)
PPCA_compound = phyl.pca(compound_tree, compound_matrix)
PPCA_raw %>% biplot(main=paste(signif(summary(PPCA_raw)$importance[2,1]*100,3),"%"), ylab=paste(signif(summary(PPCA_raw)$importance[2,2]*100,3),"%"), cex = .6, expand =1)
PPCA_compound %>% biplot(main=paste(signif(summary(PPCA_compound)$importance[2,1]*100,3),"%"), ylab=paste(signif(summary(PPCA_compound)$importance[2,2]*100,3),"%"), cex = .6, expand =0.4)
phylomorphospace(raw_tree, PPCA_raw$S)

@@

#See BAMM/BAMMsetpriors.R
#See BAMM/BAMMplots.R
#See BAMM/BAMMconfiguration.txt

## Retrieve diet data ##
#Retrieve diet info from literature BINARY
GC = read.csv("Cmerged.csv", header = T, sep=',')[,c(2,4,5,6)] %>% .[which(grepl("diet",.$character) & .$state==1),]
GC$character = factor(GC$character, levels=unique(GC$character))
GC$species = as.character(GC$species)
#Fix typos#
GC$species[which(GC$species == "Nanomia bijuga")] <- "Nanomia sp"
GC$species[which(GC$species == "Rhizophysa eyesenhardti")] <- "Rhizophysa eysenhardtii"
GC$species[which(GC$species == "Agalma okeni")] <- "Agalma okenii"
write.csv(GC, "gutcontentliteraturereview.csv")

# Retrive ROV annotation data
VARS <- read.csv("Choy2017.tsv", sep="\t")
VARS_curated = VARS[which(VARS$Pred_lowest_tax %in% ultram$tip.label | VARS$Pred_lowest_tax=="Nanomia bijuga"),]
VARS_cast = acast(VARS_curated, Pred_lowest_tax~Prey_order, fun.aggregate = length)

>>>>

```{r setup, include=FALSE, echo=FALSE, eval=FALSE}
	# Load packages
	## General
library(knitr)
library(tidyverse)
library(stringr)
library(reshape2)
	## Biological
library(Rphylip)
library(arbutus)
library(vegan)
library(ape)
library(phangorn)
library(phytools)
library(picante)
library(geiger)
library(phylobase)
library(fields)
library(phylosignal)
library(geomorph)
library(phylopath)
library(phylolm)
library(convevol)
	## Graphics
library(FactoMineR)
library(factoextra)
library(corrplot)
library(BAMMtools)
library(gridExtra)
library(xtable)
library(colorRamps)
library(ggfortify)
library(ggpubr)

@@

In [@damian2019evolution], we collected the most extensive morphological dataset on siphonophore tentilla and nematocysts using state-of-the-art microscopy techniques, and expanded the taxon sampling of the phylogeny to disentangle the evolutionary history. The analyses we carried out led to novel insights into the evolution of predatory specialization. However, this work left behind many fascinating discoveries on the evolutionary history of tentilla morphology, including nematocyst novelties, phenotypic integration, convergent evolution, paedomorphosis, and predictions on diet using morphology. Here, we present these findings and shed light on the natural history of siphonophores.

@@

We explore the evolutionary history and diversification of tentilla morphologies using phylogenetic comparative methods on the morphological dataset and phylogeny published in [@damian2019evolution].  We fitted different models generating the observed data distribution given the phylogeny for each continuous character using the function fitContinuous in the R package *geiger* [@harmon2007geiger]. These models include a non-phylogenetic whie-noise model (WN), a neutral divergence Brownian Motion model (BM), an early-burst decreasing rate model (EB), and an Ornstein-Uhlenbeck model with stabilizing selection around a fitted optimum trait value. We then ranked the models in order of increasing parametric complexity (WN, BM, EB, OU), and compared the corrected Akaike Information Criterion (AICc) support scores [@sugiura1978further] to the lowest (best) score, using a cutoff of 2 units to determine significantly better support. When the best fitting model was not significantly better than a less complex alternative, we selected the least complex model (SM13). We calculated model adequacy scores using the R package *arbutus* [@pennell2015model] (SM14). We calculated phylogenetic signals in each of the measured characters using Blomberg’s K [@blomberg2003testing] (SM13).

@@

*Evolutionary history of tentillum morphology* – [@damian2019evolution] produced the most speciose siphonophore molecular phylogeny to date, while incorporating the most recent findings in siphonophore deep node relationships. This phylogeny revealed for the first time that the genus *Erenna* is the sister to *Stephanomia amphytridis*. *Erenna* and *Stephanomia* bear the largest tentilla among all siphonophores, thus their monophyly indicates that there was a single evolutionary transition to giant tentilla. Siphonophore tentilla range in size from ~30 µm in some *Cordagalma* specimens to 2-4 cm in *Erenna* species, and up to 8 cm in *Stephanomia amphytridis* [@pugh2014review]. Most siphonophore tentilla measure between 175 and 1007 µm (1st and 3rd quartiles), with a median of 373 µm. The extreme gain of tentillum size in this newly found clade may have important implications for access to large prey size classes.

@@

  The clades defined in [@damian2019evolution] are characterized by unique evolutionary innovations in their tentilla. The clade Eucladophora (containing Pyrostephidae, Euphysonectae, and Calycophorae) encompasses all of the extant Siphonophore species (178 of 186) except Cystonects and *Apolemia*. Innovations that arose along the stem of this group include spatially segregated heteroneme and haploneme nematocysts, terminal filaments, and elastic strands (Fig. \@ref(figure7)). Pyrostephids evolved a unique bifurcation of the axial gastrovascular canal of the tentillum known as the "saccus" [@totton1965synopsis]. The stem to the clade Tendiculophora (clade containing Euphysonectae and Calycophorae) subsequently acquired further novelties such as the desmoneme and rhopaloneme (acrophore subtype ancestral) nematocysts on the terminal filament (Fig. \@ref(figure7)), which bears no other nematocyst type. These are arranged in sets of 2 parallel rhopalonemes for each single desmoneme [@skaer1988formation;@skaer1991remodelling]. The involucrum is an expansion of the epidermal layer that can cover part or all of the cnidoband (Fig. \@ref(figure2)). This structure, together with differentiated larval tentilla, appeared in the stem branch to Clade A physonects. Calycophorans evolved novelties such as larger desmonemes at the distal end of the cnidoband, pleated pedicles with a “hood” (here considered homologous to the involucrum) at the proximal end of the tentillum, anacrophore rhopalonemes, and microbasic mastigophore-type heteronemes. While calycophorans have diversified into most of the extant described siphonophore species (108 of 186), their tentilla have not undergone any major categorical gains or losses since their most recent common ancestor. Nonetheless, they have evolved a wide variation in nematocyst and cnidoband sizes.

@@

*Evolution of tentillum and nematocyst characters* – One-third of the characters measured in [@damian2019evolution] support a non-phylogenetic generative model, indicating they are not phylogenetically conserved (SM\@ref(ModelSupport)). Most (74%) characters present a significant phylogenetic signal, yet only total nematocyst volume, haploneme length, and heteroneme-to-cnidoband length ratio had a phylogenetic signal with K larger than 1. Total nematocyst volume and cnidoband-to-heteroneme length ratio showed strongly conserved phylogenetic signals. The majority (67%) of characters support BM models, indicating a history of neutral constant divergence. No relationship between phylogenetic signal and BM model support was found. Haploneme nematocyst length is the only character with support for an EB model of decreasing rate of evolution with time. No character had support for a single-optimum OU model (when uninformed by feeding guild regime priors).

@@

![\label{figure10} Phylomorphospace showing haploneme and heteroneme elongation (log scaled). Orange area delimits rod-shaped haplonemes, the blue area covers oval and round-shaped haplonemes. Smaller dots and lines represent phylogenetic relationships and ancestral states of internal nodes under BM. Species nodes in red lack either haplonemes or heteronemes, and their values are projected onto the axis of the nematocyst type they bear. Cystonects have no tentacle heteronemes and are projected onto the haploneme axis. Apolemiids have no tentacle haplonemes and are projected onto the heteroneme axis. ](~/tentilla_morph/Figures_final/figure10.jpg)

@@

*Generating dietary hypotheses using tentillum morphology* – For many siphonophore species, no feeding observations have yet been published. We generated hypotheses about the diets of these understudied siphonophores based on their known tentacle morphology and the associations between morphology and diet reported here for other species using the linear discriminant analysis of principal components (DAPC) fitted in [@damian2019evolution]. This provides concrete predictions to be tested in future work and helps extrapolate our findings to many poorly known species that are extremely difficult to collect and observe.
The discriminant analysis in [@damian2019evolution] for feeding guild (7 principal components, 4 discriminants) produced 100% discrimination, and the highest loading contributions were found for the characters (ordered from highest to lowest): Involucrum length, heteroneme volume, heteroneme number, total heteroneme volume, tentacle width, heteroneme length, total nematocyst volume, and heteroneme width. We used the predictions from this discriminant function to generate hypotheses about the feeding guild of 45 species in the morphological dataset. This extrapolation predicts that two other *Apolemia* species are gelatinous prey specialists like *Apolemia rubriversa*, and predicts that *Erenna laciniata*is a fish specialist like *Erenna richardi*. When predicting soft- and hard-bodied prey specialization, the DAPC achieved 90.9% discrimination success, only marginally confounding hard-bodied specialists with generalists (SM\@ref(DAPCsofthard)). The main characters driving this discrimination are involucrum length, heteroneme number, heteroneme volume, tentacle width, total nematocyst volume, total haploneme volume, elastic strand width, and heteroneme length. We only selected prey types with sufficient variation in the data to carry out these analyses (copepods, fish, and large crustaceans). While the presence of fish or large crustaceans in the diet cannot be unambiguously discriminated using tentillum morphology (SM\@ref(DAPCfish),SM\@ref(DAPCcrustacean)), specialization on fish or large crustacean prey can be fully disentangled (SM\@ref(DPACguilds)).

@@

*On the evolution of tentilla morphology* -- The evolutionary rate covariance results in [@damian2019evolution] indicate that tentilla are not only phenotypically integrated but also show patterns of evolutionary modularity, where different sets of characters appear to evolve in stronger correlations among each other than with other characters [@wagner1996homologues]. This may be indicative of the underlying genetic and developmental dependencies among closely homologous nematocyst types (such as desmonemes and rhopalonemes) and structures. In addition, these evolutionary modules point to hypothetical functional modules. For example, the coiling degree of the cnidoband and the extent of the involucrum have correlated rates of evolution, while high-speed videos (pers. obs.) show that the involucrum helps direct the whiplash of the uncoiling cnidoband distally (towards the prey).
The clade Tendiculophora contains far more species than its relatives Cystonectae, Apolemiidae, and Pyrostephidae. An increase in clade richness and ecological diversification can be triggered by a ‘key innovation’ [@simpson1955major]. The evolutionary innovation of the Tendiculophora tentilla with shooting cnidobands and modular regions may have facilitated further dietary diversification.

Haploneme and heteroneme shape share 21% of their variance across extant values, and 53% of the variance in their shifts along the branches of the phylogeny. However, much of this correlation is due to the contrast between Pyrostephidae and their sister group Tendiculophora. We searched for rate regime shifts in the evolution of nematocyst shape characters using a Bayesian Analysis of Macroevolutionary Mixtures (BAMM) [@rabosky2014bamm] (SM32-35). BAMM located a regime shift in heteroneme shape evolution on the branches leading to *Agalma* and *Athorybia* (SM33). For the rates of haploneme shape evolution, BAMM identified two main independent regime shifts (Fig. \@ref(figure10)): one in the branch leading to Codonophora (anisorhizas diverging from cystonects’ spherical isorhizas), and one in the branch leading to Clade B physonects (SM35). Clade B includes *Erenna*, *Stephanomia*, *Marrus*, and rhodaliids. Most of these taxa have rod-shaped anisorhizas, but *Erenna* have oval ones. No significant regime shift patterns were identified in the evolution of desmoneme and rhopaloneme shape.

@@

Our work identifies a novel example of convergent evolution. The region of the tentillum morphospace (SM4) occupied by calycophorans was independently (and more recently) occupied by the physonect *Frillagalma vityazi*. Like calycophorans, *Frillagalma* tentilla have small C-shaped cnidobands with a few rows of anisorhizas. Unlike calycophorans, they lack paired elongate microbasic mastigophores. Instead, they bear exactly three oval stenoteles, and their cnidobands are followed by a branched vesicle, unique to this genus. Their tentillum morphology is very different from that of other related physonects, which tend to have long, coiled, cnidobands with many paired oval stenoteles. Most studies on calycophoran diets have reported their prey to be primarily composed of small crustaceans, such as copepods or ostracods [@purcell1981dietary;@purcell1984functions]. The diet of *Frillagalma vityazi* is unknown, but this morphological convergence suggests that they evolved to capture similar kinds of prey. The DAPCs in [@damian2019evolution] predict that *Frillagalma* has a generalist niche with both soft and hard-bodied prey, including copepods.

@@

```{r prepdata, eval = FALSE, echo = FALSE}
# Set paths to input data
setwd("~/tentilla_morph/Organismal_paper/")

@@

## Retrieve diet data ##
GC = read.csv("~/tentilla_morph/Supplementary_materials/Dryad/literature_diet_data.tsv", header = T, sep='\t')
GC$Prey.type = factor(GC$Prey.type, levels=unique(GC$Prey.type))
GC$Siphonophore.species = as.character(GC$Siphonophore.species)
#Fix typos#
GC$Siphonophore.species[which(GC$Siphonophore.species == "Nanomia bijuga")] <- "Nanomia sp"
GC$Siphonophore.species[which(GC$Siphonophore.species == "Rhizophysa eyesenhardti")] <- "Rhizophysa eysenhardtii"
GC$Siphonophore.species[which(GC$Siphonophore.species == "Agalma okeni")] <- "Agalma okenii"
GC = GC[,1:2]
names(GC)<-c("species", "character")
#write.csv(GC, "gutcontentliteraturereview.csv")

# Retrive ROV annotation data
VARS <- read.csv("~/tentilla_morph/Supplementary_materials/Dryad/raw_ROV_data.csv")
VARS_curated = VARS[which(VARS$Siphonophore.concept %in% ultram$tip.label | VARS$Siphonophore.concept=="Nanomia bijuga"),]
VARS_cast = acast(VARS_curated, Siphonophore.concept~Prey.taxonomy, fun.aggregate = length)

hypdiet = c("Small crustacean", "Small crustacean", "Small crustacean", "Small crustacean", "Small crustacean", "Large crustacean", "Mixed", "Mixed", "Mixed", "Large crustacean", "Large crustacean", "Mixed", "Small crustacean", "Large crustacean", "Fish", "Fish", "Fish", "Large crustacean", "Gelatinous", "Fish", "Fish", "Fish")
names(hypdiet) = dprunedTree$tip.label

#Soft bodied vs hard bodied prey
soft_hard = cbind((diet$`Copepod diet`+diet$`Crustacean diet` + diet$`Decapod diet`+ diet$`Amphipod diet` + diet$`Ostracod diet`), (diet$`Fish diet`+diet$`Chaetognath diet`+diet$`Gelatinous diet`+diet$`Mollusc diet`+diet$`Polychaete diet`))
colnames(soft_hard) = c("Hard", "Soft")
rownames(soft_hard) = rownames(diet)
soft_hard <- as.data.frame(soft_hard)
soft_hard[soft_hard>0]<-1
softORhard = as.data.frame(soft_hard$Hard+soft_hard$Soft)
rownames(softORhard)=rownames(soft_hard)
names(softORhard) = "Type"
softORhard[softORhard==2]<-"Both"
softORhard[which(soft_hard$Hard==1 & soft_hard$Soft==0),]<-"Hard"
softORhard[which(soft_hard$Hard==0 & soft_hard$Soft==1),]<-"Soft"
softORhard

#PCA prep
Q_sharedmean_logs = sharedmean_logs[,-1]
#Q_sharedmean_logs = castmean_logs[,-1]
#rownames(Q_sharedmean_logs) = castmean_logs$Species
rownames(Q_sharedmean_logs) = sharedmean_logs$Species
raw_matrix = Q_sharedmean_logs[,c(1:2,4:20)] %>% .[which(!is.na(rowSums(.))),]
raw_matrix_notf = Q_sharedmean_logs[,c(1:2,4:20)] %>% .[,c(1:7,12:17)] %>% .[which(!is.na(rowSums(.))),]
raw_matrix_NaZeroes = Q_sharedmean_logs[,c(1:2,4:20)]
raw_matrix_NaZeroes[is.na(raw_matrix_NaZeroes)]<-0
compound_matrix = Q_sharedmean_logs[which(!is.na(rowSums(Q_sharedmean_logs))),c(3,21:31)]
fullraw_nozeroes <- aggregate(. ~ Species, data=castlogs[,c(-1,-3)], FUN = mean)
rownames(fullraw_nozeroes) <- fullraw_nozeroes$Species
```

```{r analyses, eval = FALSE, echo = FALSE}
### COMPARATIVE ANALYSES ###

@@

## PCA ##

#Using simple characters
#PCA(raw_matrix_NaZeroes) -> Pca_raw
PCA(raw_matrix_notf) -> Pca_raw
Pca_raw %>% fviz_contrib(choice="var", axes=1, sort.val="desc")
Pca_raw %>% fviz_contrib(choice="var", axes=2, sort.val="desc")
Pca_raw %>% fviz_pca_biplot( col.var="contrib", gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), repel = TRUE) #For figure A
Pca_raw %>% fviz_pca_biplot( col.var="contrib", gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), repel = TRUE, axes=c(3,4)) 
raw_tree = drop.tip(ultram, which(!(ultram$tip.label %in% rownames(raw_matrix_notf))))
phylomorphospace(tree=raw_tree, Pca_raw$ind$coord[,1:2], label = "horizontal", xlab = "PC1", ylab = "PC2") #For figure B
multiPhylosignal(Pca_raw$ind$coord, raw_tree)
physignal(Pca_raw$ind$coord, raw_tree)

@@

#PhyloPCA on simple character
PPCA_raw = phyl.pca(raw_tree, raw_matrix_notf)
PPCA_compound = phyl.pca(compound_tree, compound_matrix)
PPCA_raw %>% biplot(main=paste(signif(summary(PPCA_raw)$importance[2,1]*100,3),"%"), ylab=paste(signif(summary(PPCA_raw)$importance[2,2]*100,3),"%"), cex = .6, expand =1)
PPCA_compound %>% biplot(main=paste(signif(summary(PPCA_compound)$importance[2,1]*100,3),"%"), ylab=paste(signif(summary(PPCA_compound)$importance[2,2]*100,3),"%"), cex = .6, expand =0.4)
phylomorphospace(raw_tree, PPCA_raw$S, label = "horizontal")

@@

#See BAMM/BAMMsetpriors.R
#See BAMM/BAMMplots.R
#See BAMM/BAMMconfiguration.txt
```

```{r newanalyses, echo=F, eval=F}
### Morphospace analyses ###
hypfood <- data.frame(hypdiet, hypdiet, stringsAsFactors = F)
morphood <- raw_matrix_NaZeroes[which(rownames(raw_matrix_NaZeroes) %in% rownames(hypfood)),]
#morphood <- raw_matrix_notf[which(rownames(raw_matrix_notf) %in% rownames(hypfood)),]
hypfood <- hypfood[match(rownames(morphood), rownames(hypfood)),]
morphospace_data <- cbind(morphood, hypfood[,1])
#morphospace_data = cbind(raw_matrix_NaZeroes, hypfood[match(rownames(raw_matrix_NaZeroes), rownames(hypfood)),1])
names(morphospace_data)[ncol(morphospace_data)] <- "Diet"
morphospace_data$Diet <- as.character(morphospace_data$Diet) %>% str_replace_all(" ", "_")

#Test for morphospatial overlap between feeding guilds
pca_morph <- prcomp(morphospace_data[,-ncol(morphospace_data)])
autoplot(pca_morph, data = morphospace_data, frame = TRUE, label=T, colour="Diet", loadings = TRUE, loadings.colour = 'black', loadings.label = TRUE, loadings.label.size = 3, loadings.label.colour = "black")+theme_bw()
autoplot(ppp, data = morphospace_data, frame = TRUE, label=T, colour="Diet")+theme_bw()

#MANOVA for diet
LMdata <- cbind(pca_morph$x, morphospace_data$Diet) %>% as.data.frame(stringsAsFactors = F)
LMdata[,-ncol(LMdata)] <- apply(LMdata[,-ncol(LMdata)], 2, as.numeric)
names(LMdata)[ncol(LMdata)] <- "Diet"
MAN1 <- manova(cbind(PC1, PC2) ~ Diet, data = LMdata)
MAN1 %>% summary.manova() %>% .$SS %>% .$Diet -> MSS1
var_exp = (MSS1[1,2]+MSS1[2,1])/sum(MSS1)
library(geiger)
LMtree <- drop.tip(ultram, which(!(ultram$tip.label %in% rownames(LMdata))))
dat <- LMdata[,1:2]
D <- LMdata$Diet %>% as.factor()
names(D) <- rownames(LMdata)
PMAN <- aov.phylo(dat ~ D, phy = LMtree, nsim=50, test="Wilks")
PMAN  %>% summary.manova() %>% .$SS %>% .$group -> pMSS
var_exp_phyl = (pMSS[1,2]+pMSS[2,1])/sum(pMSS)

library(geomorph)
morphol.disparity(dat ~ 1, groups = D)
gdf <- geomorph.data.frame(shape = as.matrix(dat), phy = LMtree, grp = D)
ANOVA <- procD.pgls(shape ~ grp, phy = phy, data = gdf)
morphol.disparity(f1 = ANOVA, groups = ~grp, data = gdf, iter = 999)

#Fish specialization convergence
##
pcanonaphenos = PCA(raw_matrix_NaZeroes)
phenotree = drop.tip(ultram, which(!(ultram$tip.label %in% rownames(pcanonaphenos$ind$coord))))
fishtips = c(rownames(pcanonaphenos$ind$coord)[startsWith(rownames(pcanonaphenos$ind$coord), "Erenna")], "Physalia physalis", rownames(pcanonaphenos$ind$coord)[startsWith(rownames(pcanonaphenos$ind$coord), "Rhizophy")])
phylomorphospace(phenotree, pcanonaphenos$ind$coord, label = "horizontal")
convnumsig(phenotree, pcanonaphenos$ind$coord[,1:2], plot=F, convtips = fishtips, nsim=30)
convnum(phenotree, pcanonaphenos$ind$coord[,1:2], plot=T, convtips = fishtips)
convrat(phenotree,pcanonaphenos$ind$coord[,1:2], convtips = fishtips)
convratsig(phenotree,pcanonaphenos$ind$coord[,1:2], convtips = fishtips, nsim=5)
